---
title: 'Part 2: Basic Inferential Data Analysis'
author: "Sergio Paz"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Overview

Analyze the ToothGrowth data in the R datasets package.


##Load the ToothGrowth data and perform some basic exploratory data analyses

```{r}
library(datasets)
data("ToothGrowth")
head(ToothGrowth)
dim(ToothGrowth)
str(ToothGrowth)
```


##Provide a basic summary of the data

```{r}
summary(ToothGrowth)
table(ToothGrowth$supp, ToothGrowth$dose)
```


##Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose

```{r}
library(ggplot2)
library(plyr)
agg <- aggregate(len ~ dose + supp, ToothGrowth, mean)
ggplot(agg, aes(x=dose, y=len, colour = supp)) +
    geom_line(size=2, alpha=.5) + geom_point(size=5, alpha=.3) +
    xlab("Dose in miligrams") + ylab("Average Tooth length") +
    guides(colour=guide_legend(title="Supplement type")) +
    scale_color_manual(values = c("green", "#E69F01"))
```

There seems to be a correlation between the dose and tooth growth, where 
Orange Juice is more effective for lower doses, where the 2 milligram 
seems to be the maximum effect point.

```{r}
ddply(ToothGrowth, dose ~ supp,function(x) 
      c(mean=mean(x$len), sd=sd(x$len),
        conf.int=t.test(x$len)$conf.int))
```

We observe that in 95% confidence interval the `Ascorbic Acid(VC)` intervals are pairwise disjoint so we can claim with high level of confidence that the length means are distinct, moreover there is a clear growth correlation between dose & length means.

By now we can also immediately identify with high level of confidence 
that For 0.5 and 1 milligrams Orange Juice have has greater impact on 
tooth growth (On the merit that for those 2 doses there confidence
interval are pairwise disjoint).

For Orange Juice(OJ) supplement type, however, there is an overlap for dose 1 and 2 milligrams, 
and we are forced to look deeper.

```{r}
t.test(len ~ dose, paired=FALSE, var.equal=TRUE,
       data=subset(ToothGrowth, dose %in% c(1.0,2.0) & supp == 'OJ'))
```

```{r, echo=FALSE}
hp <- t.test(len ~ dose, paired=FALSE, var.equal=TRUE,
       data=subset(ToothGrowth, dose %in% c(1.0,2.0) & supp == 'OJ'))
t <- hp$statistic
p <- hp$parameter
q <- qt(.025, p)
```

The t value ```r t``` being less than **qt(.025, ```r p```)** == ```r q```
allows us to assert that the mean length for 2 milligrams as greater than the for the 1 milligram dose.

In the 2.0 milligram dose there is an overlap between `Orange Juice (OJ)` and `Ascorbic Acid (VC)` let's dig deeper 

```{r}
t.test(len ~ supp, paired=FALSE, var.equal=FALSE,
       data=subset(ToothGrowth, dose == 2.0))
```

The confidence Interval includes 0 and hence difference between the supplements types vis-a-vis mean lengths is insignificant.


##State your conclusions and the assumptions needed for your conclusions

The analysis has shown with high confidence that the there is a 
correlation between the supplement type used and teeth growth in guinea pigs, 
when for small doses of 0.5 and 1 milligrams, Orange Juice, clearly has
an advantage.