---
title: 'Part 1: Simulation Exercise'
author: "Sergio Paz"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Overview

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Here is the distribution of 1000 random uniform
```{r}
hist(runif(1000))

```

and the distribution of 1000 averages of 40 random uniforms
```{r}
mns = NULL
for(i in 1 : 1000) mns = c(mns, mean(runif(40)))
hist(mns)

```


##Simulation

```{r}
set.seed(1234)      #set the seed
lambda <- .2        #lambda
n <- 40             #sample
simulations <- 1000 #number of test
d.mean <- 1/lambda
d.sd <- 1/lambda

exp.sim <- matrix(rexp(simulations*n, rate = lambda), simulations, n)
means <- rowMeans(exp.sim)

```

###Graphic
```{r, echo=FALSE}
plot.hist.scale <- function(subject, ms, ss) {
  hist(ms, breaks=25, probability=TRUE, main=subject, xlab="", col="green")
  lines(density(ms), col="red", lwd=3) 
  curve(dnorm(x, d.mean, d.sd/sqrt(ss)), min(ms), max(ms), col="purple", add=T, lty=2, lwd=3)

  # add legend
  legend('topright', c("simulation", "theoretical"), lty=c(1,2), col=c("red", "purple"))
}  
  
plot.hist <- function(subject, ms, ss) {
  hist(ms, breaks=25, prob=TRUE,
       main=subject, xlab="", col="green")
  lines(density(ms), col="red", lwd=3) # density of the averages of samples

  abline(v=d.mean, col="darkblue", lwd=3) # Theoretical distribution center

  # theoretical density of the averages of samples
  curve(dnorm(x, d.mean, d.sd/sqrt(ss)), min(ms), max(ms), col="purple", add=T, lty=2, lwd=3)
  # add legend
  legend('topright', c("simulation", "theoretical"), lty=c(1,2), col=c("red", "purple"))
}

```

```{r, echo=FALSE}
plot.hist("Sample averages' of Exp. distribution. with rate = 0.2", means, n)
```

##Sample mean versus theoretical mean 

```{r}
sample.mean <- mean(means)
theoretical.mean <- 1/lambda
```

The deviation of the sample mean ```r sample.mean``` from the theoretical
mean ```r theoretical.mean``` is approximately 
```r round(abs(sample.mean - theoretical.mean)/theoretical.mean, 3) * 100```%

##Sample variance versus theoretical variance

```{r}
sample.sd <- sd(means)
theoretical.sd <- (d.sd)/sqrt(n)

sample.var <- var(means)
theoretical.var <- 1 / (lambda ** 2 * 40)
```

The sample variance ```r sample.var``` is different from the theoretical variance ```r theoretical.var``` by 
```r round(abs(sample.var - theoretical.var)/ theoretical.var, 3)*100```%
and the sample standard deviation ```r sample.sd``` diffrantiates from the 
theoretical standard deviation ```r theoretical.sd``` by 
```r round(abs(sample.sd - theoretical.sd)/ theoretical.sd, 3)*100```%

##Distribution

```{r}
mult = 20
par(mfrow=c(1,2))
means2 <- rowMeans(matrix(rexp(simulations*n*mult, rate=lambda), 
                 simulations, n * mult))
plot.hist("Sample averages' (40 samples)", means, n)
plot.hist("Sample averages (900 samples)", means2, n * mult)
```


###Graphic

```{r}
qqnorm(means, col="black")
qqline(means, col="red")
```
A Q-Q plot indicates as well that the 40 exponentials are quite close to the normal distribution.